---
title: "Quick look at Dr. Adeleke's 2022 Study data"
date: "`r Sys.Date()`"
author: Erin Stearns
output:
  rmdformats::downcute:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
pkgdown:
  as_is: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
#load packages
pacman::p_load(tidyverse, data.table, gfckanr,
               sf, rgdal, htmltools,
               plotly, knitr, DT, scales, kable, kableExtra,
               leaflet, leaflet.extras, tmap)
#set data directory
data_dir <- Sys.getenv("epimap_spatial")
```

```{r, include=FALSE}
#define data
url <- "https://dataexchange.gatesfoundation.org/dataset/ab6e7398-23cc-4aea-8a2c-a03e185de9e7/resource/8947389f-cb50-4a08-b8fd-6775d90281a4/download"

#load data
path <- ckan_fetch(url, store = "disk", path = tempfile())
df <- read.csv(file = path$path, header = TRUE, stringsAsFactors = F)

# ---- quick data prep

#take a look at test rows
df_test_sites <- df %>%
  filter(username %in% c("test","state_1", "training_user"))

#remove test rows
df_sans_test <- df %>%
  filter(!username %in% c("test", "state_1", "training_user"))

#separate community coordinates
df_w_coords <- df_sans_test %>%
  separate(r_GPSCommunity, sep = " ", c("latitude", "longitude", "elevation", "precision")) %>%
  #remove rows missing community coords
  filter(!is.na(longitude)) %>%
  #make new col that represents 'evidence of adult' in binary numeric form
  mutate(fly_presence = ifelse(r_EvidenceOfAdult == "yes", 1, ifelse(r_EvidenceOfAdult == "no", 0,
                                                                     -999))
         )

#make lat and long numeric
df_w_coords$latitude <- as.numeric(df_w_coords$latitude)
df_w_coords$longitude <- as.numeric(df_w_coords$longitude)

#load boundaries
adm1 <- st_read(paste0(data_dir, "/boundaries/nga/GRID3/NGA_adm1_State_Boundaries.shp"))

adm2 <- st_read(paste0(data_dir, "/boundaries/nga/GRID3/NGA_adm2_Local_Government_Area_Boundaries.shp"))
```

# Quick glance

Dr. Adeleke's study data can be found on the [GDx here.](https://dataexchange.gatesfoundation.org/dataset/ab6e7398-23cc-4aea-8a2c-a03e185de9e7/resource/8947389f-cb50-4a08-b8fd-6775d90281a4)

```{r, echo=FALSE}

pal <- colorFactor(c("green", "red"), domain = c("1", "0"))
labels <- sprintf(
  "<b>%s</b><br>%s<br>%s<br>%s",
  as.character(df_w_coords$r_State),
  paste0("LGA: ", as.character(df_w_coords$r_LGA)),
  paste0("Community name: ", as.character(df_w_coords$r_CommunityName)),
  paste0("Flies present? ", as.character(df_w_coords$r_EvidenceOfAdult))
) %>% lapply(htmltools::HTML)

leaflet(adm1) %>%
  addTiles() %>%
  addPolygons(fillOpacity = 0, 
              weight = 2,
              color = "grey") %>%
  addPolygons(data = adm2, 
              fillOpacity = 0,
              weight = 1,
              color = "grey",
              dashArray = "3") %>%
  #add sites - colored by evidence of adult flies
  addCircleMarkers(
    data = df_w_coords, ~df_w_coords$longitude, ~df_w_coords$latitude, 
    color = ~pal(df_w_coords$fly_presence), fillOpacity = 0.5,
    label = labels,
    labelOptions = labelOptions(noHide = FALSE, #textOnly = TRUE, 
                                tyle = list(
                                  "font-weight" = "normal", 
                                  padding = "3px 8px",
                                  "color" = "black"),
                                textsize = "12px",
                                direction = "auto",
                                offset = c(0,-15)),
    radius = 4
  ) %>%
  addLegend(
    "bottomright", pal = pal, 
    values = df_w_coords$fly_presence, 
    title = "Fly presence",
    opacity = 0.9
  )

```



# Summary of data by state

```{r, echo=FALSE, results='asis'}
#simplify table
df_simple <- df_w_coords %>%
  select(r_State, r_LGA, r_CommunityName,  r_EvidenceOfAdult, latitude, longitude, r_RiverName, r_RiverBasin, r_Date, r_GPS_river_basin, r_EvidenceOfLarvae, r_SiteSuitable, r_LarvaeRate, r_AdultRate)

#summarize by state
df_state <- df_simple %>%
  group_by(r_State) %>%
  summarise(sites = n())

kable(df_state, col.names = c("State", "Sites")) %>%
  kable_paper () %>%
  scroll_box(height = "300px")

```

# Data processing

Starting dataset dimensions: `r nrow(df)` rows, `r ncol(df)` columns

1. Remove 'test' rows
    + username = test, training_user, or state_1
    + Data dimensions after removing 'test' rows: `r nrow(df_sans_test)` rows
2. Prepare community coordinate data 
    + Currently a string of 4 values (lat, long, elevation, & precision), so split that into 4 individual fields
    + Transform fields to numeric fields (formerly character)
    + Remove rows missing lat/long for communities (20 rows)
    + Data dimensions after removing missing coordinates: `r nrow(df_w_coords)` rows

# Questions for Dr. Adeleke

+ There are 2 fields with coordinates: `r_GPSCommunity` & `r_GPS_river_basin` - were the coordinates from the `r_GPSCommunity` from the actual fly collection site or are these village centroids?

+ Why are some rows missing community coordinates?


# Questions for Dimagi

+ Could Dimagi please filter out 'test' rows prior to posting data?
    + Is there a consistent flag depicting a 'test row'? 




